---
title: "BEFTA bird analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(FD)
library(lme4)
library(gridExtra)
```

## R Markdown

Load data.
Note: all data and code available at https://github.com/grahamprescott/befta.birds

```{r cars}
birds.full <- read_csv("./BEFTA birds 2013-2019 final 12.06.2020.csv")
```

Structure

```{r pressure, echo=FALSE}
str(birds.full)
```

Tidy data to make sure common names are internally consistent, and consistent with functional trait dataset. Going to join datasets by common English names. Keeping the correct Latin name for Japanese tit (Parus minor) but giving it the common name 'Great Tit' for convenience in functional analysis. Check this is corrected latr. 

```{r}
birds.full %>%
  mutate(common = str_replace_all(common, c("Cinereous Tit" = "Great Tit",
                                            "Plaintive cuckoo" = "Plaintive Cuckoo",
                                            "Sunda Pied Fantail" = "Pied Fantail",
                                            "Black-naped oriole" = "Black-naped Oriole",
                                            "Pin-striped Tit-Babbler" = "Pin-striped Tit-babbler",
                                            "Striped Tit Babbler" = "Pin-striped Tit-babbler",
                                            "Pied fantail" = "Pied Fantail",
                                            "Barn swallow" = "Barn Swallow",
                                            "Edible-nest swiftlet" = "Edible-nest Swiftlet",
                                            "Slender-billed crow" = "Slender-billed Crow",
                                            "Oriental Magpie-Robin" = "Oriental Mapgie-robin",
                                            "White-breasted Kingfisher" = "White-throated Kingfisher"))) %>%
  mutate(latin = str_replace_all(latin, c("Parus cinereus" = "Parus minor",
                                          "Macronus gularis" = "Mixornis gularis",
                                          "Spilopelia chinensis" = "Stigmatopelia chinensis"))) %>%
  
  subset(uncertain == 0) %>%
  group_by(family, latin, common) %>%
  count(sort=TRUE) %>%
  view()
```

Adjust dataset

```{r}
birds.adjusted <- birds.full %>%
  mutate(common = str_replace_all(common, c("Cinereous Tit" = "Great Tit",
                                            "Plaintive cuckoo" = "Plaintive Cuckoo",
                                            "Sunda Pied Fantail" = "Pied Fantail",
                                            "Black-naped oriole" = "Black-naped Oriole",
                                            "Pin-striped Tit-Babbler" = "Pin-striped Tit-babbler",
                                            "Striped Tit Babbler" = "Pin-striped Tit-babbler",
                                            "Pied fantail" = "Pied Fantail",
                                            "Barn swallow" = "Barn Swallow",
                                            "Edible-nest swiftlet" = "Edible-nest Swiftlet",
                                            "Slender-billed crow" = "Slender-billed Crow",
                                            "Oriental Magpie-Robin" = "Oriental Mapgie-robin",
                                            "White-breasted Kingfisher" = "White-throated Kingfisher"))) %>%
  mutate(latin = str_replace_all(latin, c("Parus cinereus" = "Parus minor",
                                          "Macronus gularis" = "Mixornis gularis",
                                          "Spilopelia chinensis" = "Stigmatopelia chinensis")))
```

Check uncertain records (tiny fraction)

```{r}
birds.full %>%
  subset(uncertain != 0) %>%
  view()
```

Create 'birds.core' for the core plot analysis

```{r}
birds.core <- birds.adjusted %>%
  # remove flyovers and observations > 70 m [to only include records in core plot]
  subset(distance < 30) %>%
  # remove records of uncertain ID [doubtful records, those marked 'Unidentified', Sunda frogmouth etc had value of 1 in 'uncertain' column]
  subset(uncertain == 0) %>%
  # remove red junglefowl / domestic chicken
  subset(common != "Red Junglefowl") %>%
  #remove Eurasian tree sparrow
  subset(common != "Eurasian Tree Sparrow") %>%
  # remove Oriental Mapgie Robin (poaching pressure biases results)
  subset(common != "Oriental Magpie-robin") %>%
  # remove Sunda Frogmout (nocturnal birds not adequately sampled)
  subset(common != "Sunda Frogmouth")
```

Visualise change in species richness (core dataset)

```{r}
birds.core %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_boxplot(aes(x = stage, y = sr, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,5) +
  facet_wrap(~treatment)  +
  labs( y = "number of species recorded", x = "", title = "")

ggsave(filename = paste0("species.richness.effect",format(Sys.time(), "%d-%b-%Y %H.%M"), ".pdf"), width = 140, height = 160, dpi = "print", units = "mm", device='pdf')
```

Abundance patterns for core area?

```{r}
birds.core %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_boxplot(aes(x = stage, y = n, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  facet_wrap(~treatment)  +
  labs( y = "number of individuals recorded", x = "", title = "")

ggsave(filename = paste0("abundance.effect",format(Sys.time(), "%d-%b-%Y %H.%M"), ".pdf"), width = 140, height = 160, dpi = "print", units = "mm", device='pdf')
```

Load vegetation data

```{r}
veg.all <- read_csv("BEFTA_InsectTrapVeg-RReady.csv")
```

```{r}
head(veg.all)
```
subset

```{r}
veg <- veg.all %>%
  filter(TimeBlock == "2013-03" | TimeBlock == "2017-05") %>%
  select(PrePost, TimeBlock, TrapBearing, BEFTAcode, PlotNumber, Treatment, VegHeight, BareGround,
         Fern, Frond, OtherPlants, Densio1, Densio2, Densio3, Densio4, MeanPercentOpen)

head(veg)
```
Need to ask Sarah about the variables, e.g. what is TrapBearing https://www.frontiersin.org/articles/10.3389/ffgc.2019.00033/full

Below is the kind of code I can use to prep predictor variables (i.e. averaging across readings). But need to check I know what I actually want!! 
```{r}
 veg %>%
  unite("StagePlot", c(PrePost, PlotNumber), sep = "_") %>%
  group_by(StagePlot) %>%
  summarise(mean.VegHeight = mean(VegHeight), 
            mean.Fern = mean(Fern),
            mean.BareGround = mean(BareGround),
            mean.Frond = mean(Frond)) %>%
  ggplot() +
  geom_point(aes(x = StagePlot, y = mean.VegHeight))

```



```{r}
veg.vars <- veg %>%
  unite("StagePlot", c(PrePost, PlotNumber), sep = "_") %>%
  group_by(StagePlot) %>%
  summarise(mean.VegHeight = mean(VegHeight), 
            mean.Fern = mean(Fern),
            mean.BareGround = mean(BareGround),
            mean.Frond = mean(Frond)) %>%
  separate(StagePlot, into = c("Stage", "Plot"), sep = "_") %>%
  mutate(Stage = str_replace_all(Stage, c("1-Pre" = "Before",
                                          "2-Post" = "After"))) %>%
  rename(stage = Stage) %>%
  rename(plot = Plot)

```


Combine veg.vars with birds.core

```{r}
combo.data <- full_join(veg.vars, birds.core)
view(combo.data)
```

Preliminary graphing

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.VegHeight, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.Frond, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.Fern, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```
```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.BareGround, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.VegHeight, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean vegetation height", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.VegHeight, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean vegetation height", title = "")
```


```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Frond, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean frond", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Frond, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean frond", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Fern, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean fern", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Fern, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean fern", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.BareGround, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean bare ground", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.BareGround, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean bare ground", title = "")
```


FUNCTIONAL DIVERSITY METRICS

Going to start the analysis with the 'birds.core' file. 

Load in the functional traits data (from Elton Traits)

```{r}
functional.traits <- read.csv("./functional.traits.csv")
names(functional.traits)
```

Need to make sure Latin names are aligned


```{r}
unique(birds.core$latin)
```

vs 

```{r}
unique(functional.traits$Scientific)
view(functional.traits)
```


```{r}

```


END