---
title: "BEFTA bird analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(FD)
library(lme4)
library(gridExtra)
```

## R Markdown

Load data. 
Tidy data to make sure common names are internally consistent, and consistent with functional trait dataset. Going to join datasets by common English names. Note, Japanese tit as 'Great Tit' / 'Parus major' for convenience in functional analysis.  

Note: all data and code available at https://github.com/grahamprescott/befta.birds

```{r cars}
birds.adjusted <- read_csv("./BEFTA birds 2013-2019 final 12.06.2020.csv") %>%
  mutate(common = str_replace_all(common, c("Cinereous Tit" = "Great Tit",
                                            "Plaintive cuckoo" = "Plaintive Cuckoo",
                                            "Sunda Pied Fantail" = "Pied Fantail",
                                            "Black-naped oriole" = "Black-naped Oriole",
                                            "Pin-striped Tit-Babbler" = "Pin-striped Tit-babbler",
                                            "Striped Tit Babbler" = "Pin-striped Tit-babbler",
                                            "Pied fantail" = "Pied Fantail",
                                            "Barn swallow" = "Barn Swallow",
                                            "Edible-nest swiftlet" = "Edible-nest Swiftlet",
                                            "Slender-billed crow" = "Slender-billed Crow",
                                            "Oriental Magpie-Robin" = "Oriental Mapgie-robin",
                                            "White-breasted Kingfisher" = "White-throated Kingfisher"))) %>%
  mutate(latin = str_replace_all(latin, c("Parus cinereus" = "Parus major",
                                          "Macronus gularis" = "Mixornis gularis",
                                          "Spilopelia chinensis" = "Stigmatopelia chinensis")))
```

Create 'birds.core' for the core plot analysis. Specifically, remove flyovers, only include observations <30 m from observer (to be sure that they are in the plot), and remove some problem species: red jungle fowl (confusion with domestic chickens), Eurasian tree sparrow (associated with settlements not core oil palm), oriental magpie robin (not enough records, trapping concern), Sunda frogmouth (nocturnal)

```{r}
birds.core <- birds.adjusted %>%
  # remove flyovers and observations > 70 m [to only include records in core plot]
  subset(distance < 30) %>%
  # remove records of uncertain ID [doubtful records, those marked 'Unidentified', Sunda frogmouth etc had value of 1 in 'uncertain' column]
  subset(uncertain == 0) %>%
  # remove red junglefowl / domestic chicken
  subset(common != "Red Junglefowl") %>%
  #remove Eurasian tree sparrow
  subset(common != "Eurasian Tree Sparrow") %>%
  # remove Oriental Mapgie Robin (poaching pressure biases results)
  subset(common != "Oriental Magpie-robin") %>%
  # remove Sunda Frogmout (nocturnal birds not adequately sampled)
  subset(common != "Sunda Frogmouth")
```

Visualise change in species richness (core dataset)

```{r}
birds.core %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_boxplot(aes(x = stage, y = sr, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,5) +
  facet_wrap(~treatment)  +
  labs( y = "number of species recorded", x = "", title = "")

ggsave(filename = paste0("species.richness.effect",format(Sys.time(), "%d-%b-%Y %H.%M"), ".pdf"), width = 140, height = 160, dpi = "print", units = "mm", device='pdf')
```

Abundance patterns for core area?

```{r}
birds.core %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_boxplot(aes(x = stage, y = n, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  facet_wrap(~treatment)  +
  labs( y = "number of individuals recorded", x = "", title = "")

ggsave(filename = paste0("abundance.effect",format(Sys.time(), "%d-%b-%Y %H.%M"), ".pdf"), width = 140, height = 160, dpi = "print", units = "mm", device='pdf')
```


FUNCTIONAL DIVERSITY METRICS

Going to start the analysis with the 'birds.core' file. 

Load in the functional traits data (from Elton Traits)

```{r}
functional.traits <- read.csv("./functional.traits.csv")
names(functional.traits)
```

Need to make sure Latin names are aligned


```{r}
unique(birds.core$latin)
```

vs 

```{r}
unique(functional.traits$Scientific)
```

Check complete. Now onto the functional diversity analysis! 

Note: need to mutate functional.traits before I join - e.g. Scientific -> latin; English -> common

Next step is to merge the datasets. 

```{r}
functional.diversity.analysis <-  full_join(functional.traits, birds.core)
view(combo.data)
```


END