---
title: "R Notebook"
output: html_notebook
---

R markdown file for exploring the vegetation variables

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(FD)
library(lme4)
library(gridExtra)
```

These will be used as the predictor variables in models, to see if they improve predictions above treatment. This will also enable us to visualise the effects of treatment. 

Load bird data for comparison later:

```{r cars}
birds.full <- read_csv("./BEFTA birds 2013-2019 final 12.06.2020.csv")
```

```{r}
birds.adjusted <- birds.full %>%
  mutate(common = str_replace_all(common, c("Cinereous Tit" = "Great Tit",
                                            "Plaintive cuckoo" = "Plaintive Cuckoo",
                                            "Sunda Pied Fantail" = "Pied Fantail",
                                            "Black-naped oriole" = "Black-naped Oriole",
                                            "Pin-striped Tit-Babbler" = "Pin-striped Tit-babbler",
                                            "Striped Tit Babbler" = "Pin-striped Tit-babbler",
                                            "Pied fantail" = "Pied Fantail",
                                            "Barn swallow" = "Barn Swallow",
                                            "Edible-nest swiftlet" = "Edible-nest Swiftlet",
                                            "Slender-billed crow" = "Slender-billed Crow",
                                            "Oriental Magpie-Robin" = "Oriental Mapgie-robin",
                                            "White-breasted Kingfisher" = "White-throated Kingfisher"))) %>%
  mutate(latin = str_replace_all(latin, c("Parus cinereus" = "Parus major",
                                          "Macronus gularis" = "Mixornis gularis",
                                          "Spilopelia chinensis" = "Stigmatopelia chinensis")))
```

```{r}
birds.core <- birds.adjusted %>%
  # remove flyovers and observations > 70 m [to only include records in core plot]
  subset(distance < 30) %>%
  # remove records of uncertain ID [doubtful records, those marked 'Unidentified', Sunda frogmouth etc had value of 1 in 'uncertain' column]
  subset(uncertain == 0) %>%
  # remove red junglefowl / domestic chicken
  subset(common != "Red Junglefowl") %>%
  #remove Eurasian tree sparrow
  subset(common != "Eurasian Tree Sparrow") %>%
  # remove Oriental Mapgie Robin (poaching pressure biases results)
  subset(common != "Oriental Magpie-robin") %>%
  # remove Sunda Frogmout (nocturnal birds not adequately sampled)
  subset(common != "Sunda Frogmouth")
```


Load vegetation data

```{r}
veg.all <- read_csv("BEFTA_InsectTrapVeg-RReady.csv")
```

```{r}
head(veg.all)
```
subset

```{r}
veg <- veg.all %>%
  filter(TimeBlock == "2013-03" | TimeBlock == "2017-05") %>%
  select(PrePost, TimeBlock, TrapBearing, BEFTAcode, PlotNumber, Treatment, VegHeight, BareGround,
         Fern, Frond, OtherPlants, Densio1, Densio2, Densio3, Densio4, MeanPercentOpen)

head(veg)
```
Need to ask Sarah about the variables, e.g. what is TrapBearing https://www.frontiersin.org/articles/10.3389/ffgc.2019.00033/full

Below is the kind of code I can use to prep predictor variables (i.e. averaging across readings). But need to check I know what I actually want!! 
```{r}
 veg %>%
  unite("StagePlot", c(PrePost, PlotNumber), sep = "_") %>%
  group_by(StagePlot) %>%
  summarise(mean.VegHeight = mean(VegHeight), 
            mean.Fern = mean(Fern),
            mean.BareGround = mean(BareGround),
            mean.Frond = mean(Frond)) %>%
  ggplot() +
  geom_point(aes(x = StagePlot, y = mean.VegHeight))

```



```{r}
veg.vars <- veg %>%
  unite("StagePlot", c(PrePost, PlotNumber), sep = "_") %>%
  group_by(StagePlot) %>%
  summarise(mean.VegHeight = mean(VegHeight), 
            mean.Fern = mean(Fern),
            mean.BareGround = mean(BareGround),
            mean.Frond = mean(Frond)) %>%
  separate(StagePlot, into = c("Stage", "Plot"), sep = "_") %>%
  mutate(Stage = str_replace_all(Stage, c("1-Pre" = "Before",
                                          "2-Post" = "After"))) %>%
  rename(stage = Stage) %>%
  rename(plot = Plot)

```


Combine veg.vars with birds.core

```{r}
combo.data <- full_join(veg.vars, birds.core)
view(combo.data)
```

Preliminary graphing

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.VegHeight, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.Frond, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.Fern, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```
```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  ggplot() +
  geom_boxplot(aes(x = treatment, y = mean.BareGround, fill = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  facet_wrap(~stage) +
  theme_bw(12)
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.VegHeight, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean vegetation height", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.VegHeight, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean vegetation height", title = "")
```


```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Frond, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean frond", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Frond, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean frond", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Fern, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean fern", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.Fern, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean fern", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.BareGround, y = n, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of individuals recorded", x = "mean bare ground", title = "")
```

```{r}
combo.data %>%
  mutate(stage = fct_relevel(stage, "Before", "After")) %>%
  group_by(round, plot) %>%
  mutate(n = length(common)) %>%
  mutate(sr = n_distinct(common)) %>%
  ggplot() +
  geom_point(aes(x = mean.BareGround, y = sr, color = treatment)) +
  scale_fill_manual(values = c("#1b9e77", "#7570b3", "#d95f02")) +
  theme_bw(12) +
  ylim(0,8) +
  #facet_wrap(~stage) +
  labs( y = "number of species recorded", x = "mean bare ground", title = "")
```

